---
title: "Preference Allocation"
author: "Avision Ho"
date: "9 June 2018"
output: 
  ioslides_presentation:
    widescreen: true
    transition: faster
    mathjax: default
    keep_md: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Load packages
#install.packages('markdown')
library(matchingR)
library(readr)
library(dplyr)
library(tibble)
library(DT)
library(pander)

# Set random number generator so so results are replicable
set.seed(1)

# Load data
utility_delegates <- read_csv(file = "data/dummy_student_preferences.csv")

# Import source code
source("scripts/main.R")
```

## Author's Note

- **DISCLAIMER:** The author is relatively new to the **Gale-Shapley algorithm** so would appreciate suggestions and tips.
- The **Gale-Shapley algorithm** is unlikely to be the most optimal solution to the problem of allocating delegates with preferences to sessions. 
    - With additional time, the author wishes to explore an algorithm that maximises the group utility of delegates.


## Preference Allocation | One-sided matching of delegates to sessions

> - We have a *n* delegates who have preferences over *m* sessions they wish to go to. 
> - Delegates can only go to one session at any one time. 
> - Each session can only hold varying numbers of people.

# Or more formally...

## Defining the Problem {.smaller}

> - Each delegate, $d_i$ for $i = 1, ..., n$, receives utility $u_{d_i}(s_j)$ from being matched to session <b> $s_j$ </b>
> - Each session, $s_j$ for $j = 1, ..., m$, receives utility $u_{s_j}(d_i)$ = 0 from being matched to delegate $d_i$. [^1]

[^1]: This means sessions are indifferent over which delegates attends.

> - Each delegate, $d_i$, has a well defined preference ordering over all sessions. 

<div class="centered"; style="color:blue;">
For example, for delegate $d_i$ who has the choice of two sessions $s$ and $s'$, they prefer session $s$ over session $s'$ when $u_{d_i}(s) \ge u_{d_i}(s')$
</div>

> - Each session, $s_j$, can hold $p_k$ delegates where $k \in \{1, ..., m\}$. 

<p style="font-size:10px;">
Note that $i$, $j$ and $k$ belong to the set of natural numbers. 
</p>

## Background | Nobel Prize for Economics 2012 {.smaller}

> - The issue we will tackle in this project is popularised by the famous **stable marriage problem**.
> - At a high level, it asks the question of how a set of men and women should be matched whilst respecting each individual's preferences. 
> -  Gale and Shapley first analysed this matching at an abstract, general level in 1962, and proved that for any equal number of men and women, it is always possible to solve the **stable marriage problem** and make all marriages stable.
> -  The Nobel Prize in Economic Sciences 2012 was awared to Lloyd Shapely and Alvin Roth for extending this work in areas such as assigning new doctors to hospitals, students to colleges, and human organ transplants to recipients. [^2]

[^2]: The [Sveriges Riksbank prize in Economic Sciences in Memory of Alfred Nobel 2012](https://www.nobelprize.org/nobel_prizes/economic-sciences/laureates/2012/popular-economicsciences2012.pdf) was the introductory lecture your author received when starting his first year of BSc. Mathematics and Economics course at the London School of Economics and Political Science (LSE). 

## Key Assumptions {.smaller}
Key assumptions to our problem and the **stable marriage problem** are:

> -  Each man (*delegate*) and woman (*session*) has a well-defined preference ordering, meaning we know their preferences.
> -  Each man (*delegate*) and woman (*session*) prefers being matched over being not matched. 

Our problem differs from the **stable marriage problem** by:

> - Sessions have no preferences over which delegates attend them. We thus have one-sided matching.
> - *Stability* here will be defined from the viewpoint of *pareto efficiency*. [^3]

[^3]: A delegate, $d_i$, may gain more utility from attending a session, $s'$, different to the one they were matched with, $s$. However, this will come at the cost of reducing the utility of the delegate, $d_i'$ who was matched to the session the initial delegate, $d_i$, wanted to attend. $u_{d_i}(s') >  u_{d_i}(s)$ but $u_{d_i'}(s) < u_{d_i'}(s')$  

> - Each session can hold a varying number of people.

## How it Works {.smaller}
The **Gale-Shapley algorithm** for the **college admissions problem** will be applied for our problem. It involves a number of rounds:

  (@) Students sequentially make proposals to each off their most preferred available colleges.
  (@) A college can hold onto at most $s$ proposals at a time, since they have $s$ spaces for students.
  (@) A college with an open slot will provisionally accept any application it receives.
  (@) A college already holding $s$ applications will reject any student application that it values less than the current applicants they have provisionally accepted.
  (@) A college already holding $s$ applications will accept any student application that it values more than the current applicants they have provisionally accepted, and drop the least valued applicant they have currently provisionally accepted.
  (@) Process continues until all students are matched to colleges.


## Data
- For illustrative purposes, we will use dummy data to model preferences. The table below is of *delegate preferences*.

- For instance, we can see that for *Person 1*, they strictly prefer *Session 2* over *Session 1*, $u_{d_1}(2) > u_{d_1}(1)$.

```{r Data: delegate preferences}
utility_delegates %>% 
  datatable()
```

***

- The *session preferences* table is the transpose of the *delegate preferences* table above.

- We have given each session a preference of $0$ for each delegate. This is because sessions are indifferent over which delegates attend. 

```{r Data: session preferences}
utility_sessions %>% 
  datatable()
```

***

- Suppose that `Session 1` and `Session 4` can hold two delegates, whereas `Session 2` and `Session 3` can only hold one delegate. 
- This means that in total, all sessions can hold 6 delegates.
- Given we have 6 delegates to allocate to sessions, we will not get any unmatched delegates to sessions.

## Analysis {.smaller}
In the code below, we will run the college admissions variant of the **Gale-Shapley algorithm**.

``` {r Gale-Shapley Analysis, echo = TRUE, eval = FALSE}
galeShapley.collegeAdmissions(studentUtils = utility_delegates, 
                              collegeUtils = utility_sessions, 
                              slots = c(2, 1, 1, 2))
```

From running the algorithm we have the following allocation of delegates to sessions. `r results$matched.students %>% t()`. 

<div class="centered"; style="color:blue;">
For instance, `Person 1` and `Person 4` are allocated to `r paste0("Session ", results$matched.students[1,])` and `r paste0("Session ", results$matched.students[4,])` respectively. 
</div>

## Algorithm's Performance {.smaller}
We will present the matched allocations generated by the college admissions variant of the **Gale-Shapley algorithm** we used against the initial preferences of delegates to check that our matchings are desirable,

```{r Compare Outputs with Preferences}
names <- c("Person 1", "Person 2", "Person 3", "Person 4", "Person 5", "Person 6")
rownames(results$matched.students) <- names 
colnames(results$matched.students) <- "Session Allocated"
results$matched.students %>% 
  t()

utility_delegates %>% 
  datatable()
```

# Hope you enjoyed the talk! :)
