---
title: "Preference Allocation"
author: "Avision Ho"
date: "9 June 2018"
output: 
  ioslides_presentation:
    widescreen: true
    transition: faster
    mathjax: default
    keep_md: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Load packages
library(matchingR)
library(readr)
library(dplyr)
library(tidyr)
library(tibble)
library(DT)

# Set random number generator so so results are replicable
set.seed(1)

# Load data - note are moving up one directory
utility_delegates <- read_csv(file = "../data/dummy_student_preferences.csv")

# Load functions
source("../src/functions.R")
```


```{r data-prep, include=FALSE}
utility_delegates <- utility_delegates %>% 
  remove_rownames() %>% 
  column_to_rownames(var = "X1")

# Set colleges to have no preferences
n_delegates <- ncol(utility_delegates)
m_sessions <- nrow(utility_delegates)

utility_sessions <- matrix(data = rep(x = 0, times = n_delegates*m_sessions),
                           nrow = n_delegates, 
                           ncol = m_sessions)
utility_sessions <- utility_sessions %>% 
  as_tibble() %>% 
  rename(`College 1` = V1,
         `College 2` = V2,
         `College 3` = V3,
         `College 4` = V4)
```

## Author's Note

- **DISCLAIMER:** The author is relatively new to the **Gale-Shapley algorithm** so would appreciate suggestions and tips.
- The **Gale-Shapley algorithm** is unlikely to be the most optimal solution to the problem of allocating delegates with preferences to sessions. 
    - With additional time, the author wishes to explore an algorithm that maximises the group utility of delegates.
    
- Credit to a Fatma Hussain, a work experience student who came up with the second algorithm, **iterative preferences**.


## Preference Allocation | One-sided matching of delegates to sessions

> - We have a *n* delegates who have preferences over *m* sessions they wish to go to. 
> - Delegates can only go to one session at any one time. 
> - Each session can only hold varying numbers of people.

# Or more formally...

## Defining the Problem {.smaller}

> - Each delegate, $d_i$ for $i = 1, ..., n$, receives utility $u_{d_i}(s_j)$ from being matched to session <b> $s_j$ </b>
> - Each session, $s_j$ for $j = 1, ..., m$, receives utility $u_{s_j}(d_i)$ = 0 from being matched to delegate $d_i$. [^1]

[^1]: This means sessions are indifferent over which delegates attends.

> - Each delegate, $d_i$, has a well defined preference ordering over all sessions. 

<div class="centered"; style="color:blue;">
For example, for delegate $d_i$ who has the choice of two sessions $s$ and $s'$, they prefer session $s$ over session $s'$ when $u_{d_i}(s) \ge u_{d_i}(s')$
</div>

> - Each session, $s_j$, can hold $p_k$ delegates where $k \in \{1, ..., m\}$. 

<p style="font-size:15px;">
Note that $i$, $j$ and $k$ belong to the set of natural numbers. 
</p>

## Approaches
To tackle this preference allocation problem, we will take two alternative approaches:

    - Gale-Shapley algorithm
    - Iterative Preference algorithm

## Background | Nobel Prize for Economics 2012 {.smaller}

> - The issue we will tackle in this project is popularised by the famous **stable marriage problem**.
> - At a high level, it asks the question of how a set of men and women should be matched whilst respecting each individual's preferences. 
> -  Gale and Shapley first analysed this matching at an abstract, general level in 1962, and proved that for any equal number of men and women, it is always possible to solve the **stable marriage problem** and make all marriages stable.
> -  The Nobel Prize in Economic Sciences 2012 was awared to Lloyd Shapely and Alvin Roth for extending this work in areas such as assigning new doctors to hospitals, students to colleges, and human organ transplants to recipients. [^2]

[^2]: The [Sveriges Riksbank prize in Economic Sciences in Memory of Alfred Nobel 2012](https://www.nobelprize.org/nobel_prizes/economic-sciences/laureates/2012/popular-economicsciences2012.pdf) was the introductory lecture your author received when starting his first year of BSc. Mathematics and Economics course at the London School of Economics and Political Science (LSE). 

## Key Assumptions {.smaller}
Key assumptions to our problem and the **stable marriage problem** are:

> -  Each man (*delegate*) and woman (*session*) has a well-defined preference ordering, meaning we know their preferences.
> -  Each man (*delegate*) and woman (*session*) prefers being matched over being not matched. 

Our problem differs from the **stable marriage problem** by:

> - Sessions have no preferences over which delegates attend them. We thus have one-sided matching.
> - *Stability* here will be defined from the viewpoint of *pareto efficiency*. [^3]

[^3]: A delegate, $d_i$, may gain more utility from attending a session, $s'$, different to the one they were matched with, $s$. However, this will come at the cost of reducing the utility of the delegate, $d_i'$ who was matched to the session the initial delegate, $d_i$, wanted to attend. $u_{d_i}(s') >  u_{d_i}(s)$ but $u_{d_i'}(s) < u_{d_i'}(s')$  

> - Each session can hold a varying number of people.

## How it Works | Gale-Shapley algorithm {.smaller}
The **Gale-Shapley algorithm** for the **college admissions problem** involves a number of rounds:

  (@) Students sequentially make proposals to each off their most preferred available colleges.
  (@) A college can hold onto at most $s$ proposals at a time, since they have $s$ spaces for students.
  (@) A college with an open slot will provisionally accept any application it receives.
  (@) A college already holding $s$ applications will reject any student application that it values less than the current applicants they have provisionally accepted.
  (@) A college already holding $s$ applications will accept any student application that it values more than the current applicants they have provisionally accepted, and drop the least valued applicant they have currently provisionally accepted.
  (@) Process continues until all students are matched to colleges.

## How it Works | Iterative Preference algorithm
The **Iterative Preference algorithm** involves a number of rounds:

  (@) Pick a delegate at random and assign them to their most preferred session.
  (@) If this session is full, then assign them to their next most preferred session.
  (@) Repeat the previous step until this randomly selected delegate is allocated to a free session.
  (@) Pick another delegate  at random and repeat the previous steps until they are allocated to a free session.
  (@) Process continues until all delegates have been assigned to a free session.

## Data
- For illustrative purposes, we will use dummy data to model preferences. The table below is of *delegate preferences*.

- For instance, we can see that for *Person 1*, they strictly prefer *Session 2* over *Session 1*, $u_{d_1}(2) > u_{d_1}(1)$.

```{r Data: delegate preferences}
datatable(data = utility_delegates)
```

***

- The *session preferences* table is the transpose of the *delegate preferences* table above.

- We have given each session a preference of $0$ for each delegate. This is because sessions are indifferent over which delegates attend. 

```{r Data: session preferences}
datatable(data = utility_sessions)
```

***

- Suppose that `Session 1` and `Session 4` can hold two delegates, whereas `Session 2` and `Session 3` can only hold one delegate. 
- This means that in total, all sessions can hold 6 delegates.
- Given we have 6 delegates to allocate to sessions, we will not get any unmatched delegates to sessions.

## Analysis Approach 1 - Gale-Shapley {.smaller}
In the code below, we will run the college admissions variant of the **Gale-Shapley algorithm**.

``` {r Gale-Shapley Analysis, echo = TRUE}
results_galeshapley <- galeShapley.collegeAdmissions(
  studentUtils = utility_delegates,
  collegeUtils = utility_sessions,
  slots = c(2, 1, 1, 2)
)
```

From running the algorithm we have the following allocation of delegates to sessions. `r t(results_galeshapley$matched.students)`. 

<div class="centered"; style="color:blue;">
For instance, `Person 1` and `Person 4` are allocated to `r paste0("Session ", results_galeshapley$matched.students[1,])` and `r paste0("Session ", results_galeshapley$matched.students[4,])` respectively. 
</div>

## Algorithm's Output {.smaller}
In the first output is the matched allocations from using the **Gale-Shapley algorithm**.

```{r Gale-Shapley - allocation}
# convert to viewable format for slides
names <- c("Person 1", "Person 2", "Person 3", "Person 4", "Person 5", "Person 6")
rownames(results_galeshapley$matched.students) <- names 
colnames(results_galeshapley$matched.students) <- "Session Allocated"

# show allocation
t(results_galeshapley$matched.students) 
```

In the second output is the table of delegates' initial preferences to compare our allocations against.

```{r Gale-Shapley - delegate preferences}
# show delegate preferences
datatable(data = utility_delegates)
```


## Analysis Approach 2 - Iterative Preferences {.smaller}
In the code below, we will run Fatma's suggested algorithm, **Iterative Preferences** against the initial preferences of delegates to check that our matchings are desirable.

Before we do so, we need to reshape our data into tidy data format.
```{r Iterative Preference - wrangle, include=FALSE}
utility_delegates <- utility_delegates %>% 
  rownames_to_column(var = "Session") %>% 
  pivot_longer(cols = -"Session", 
               names_to = "Person",
               values_to = "Preference Score") %>% 
  pivot_wider(id_cols = "Person",
              names_from = "Session",
              values_from = "Preference Score")
room_sizes <- data.frame(Room = c("Room_01","Room_02","Room_03","Room_04"),
                         Size = c(2, 1 , 1, 2))

```


``` {r Iterative Preference - implementation, echo = TRUE}
results_iterativepreference <- func_iterative_preferences(x = utility_delegates, 
                                                          limits = room_sizes, 
                                                          with_replacement = FALSE
)
```

```{r Iterative Preference - helper, include = FALSE}
matchings <- results_iterativepreference[[1]]
sampling <- results_iterativepreference[[2]]
```


From running the algorithm we have the following allocation of delegates to sessions. `r matchings$SessionPreferredColumnId`. 

<div class="centered"; style="color:blue;">
For instance, `r pull(matchings[1,1])` and `r pull(matchings[4,1])` are allocated to `r paste0("Session ", pull(matchings[1,2]))` and `r paste0("Session ", pull(matchings[4,2]))` respectively. 
</div>

## Algorithm's Output {.smaller}
In the first output is the matched allocations from using the **Iterative Preferences algorithm**. 

```{r Iterative Preference - allocation}
# show allocation
matchings <- matchings %>% 
  rename("Session Allocated" = "SessionPreferredColumnId") %>% 
  mutate(PersonRowId = paste0('Person ', PersonRowId)) %>% 
  t() %>% 
  as_tibble()
colnames(matchings) <- matchings[1, ]
matchings <- matchings[-1, ]
datatable(data = matchings,
          options = list(dom = 't'),
          rownames = FALSE)
```

In the second output is the random order of delegates that our algorithm used. 

```{r Iterative Preference - delegate sampling}
# show delegate sampling
paste0("Person ", sampling)
```

In the third output is the table of delegates' initial preferences to compare our allocations against.

```{r Iterative Preference - delegate preferences}
# show delegate preferences
datatable(data = utility_delegates,
          options = list(dom = 't'),
          rownames = FALSE)
```

# Hope you enjoyed the talk! :)
